---
title: Final Lipidomic Analysis
author: "Andrew Patt"
date: "June 17th, 2020"
output:
  html_document:
      toc: true
      toc_float:
        collapsed: false
      theme: cerulean
      highlight: tango
---

```{r,echo=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, 
    fig.height=6,fig.width = 12,fig.fullwidth = TRUE)
```

The code in this script was used to identify lipidomic differences between MDM2 higher and lower cell lines, as well as differences driven by atorvastatin treatment. 

# Install and load required packages

```{r, Libraries}
library(tidyverse)
library(RColorBrewer)
library(gplots)
library(readxl)
library(plotly)
library(reshape)
library(lme4)
library(pbapply)
library(sjstats)
library(UpSetR)
library(RColorBrewer)
library(cowplot)
```

```{r, Set statistical cutoffs}
p_thresh <- 0.05
fc_thresh <- 0.75
```

## Load data and metadata
```{r, Set up statistical groups}
load("NewSarcomaLines_lipidomics.Rda")

lipidomic_key<-read.csv("Lipidomic_Key_Updated.csv")
lipidomic_key <- lipidomic_key[match(rownames(lipidDDLPS),lipidomic_key$name),]

```

These data were already normalized and imputed in an external script,
so we can move straight to statistical analysis.

## Panel composition
```{r, Panel composition, fig.width=18, fig.height=8}

palette<-c(brewer.pal(12,"Paired"),"#40E0D0","violet","grey80")

lipidomic_key <- within(lipidomic_key, 
                   LM.Main.Class <- factor(LM.Main.Class, 
                                      levels=names(sort(table(LM.Main.Class), 
                                                        decreasing=TRUE))))

ggplot(lipidomic_key,aes(x=LM.Main.Class,fill=LM.Main.Class))+
    geom_bar()+
    labs(y = "# of Lipids") +
    theme_classic(base_size=22) +
    theme(axis.text.x = element_text(angle=30,
                                     hjust=1,size=22,
                                     face="bold"),
          axis.text.y =element_text(size=36),
          plot.margin=unit(c(5.5, 5.5, 5.5, 150), "points")) +
    scale_fill_manual(values=palette) +
    labs(x="") +
    guides(fill=FALSE)
    
```

# Linear Mixed Effect Modeling

## All samples {.tabset}
### Linear Mixed effect models (p value histograms)
#### Atorvastatin Treated vs. Untreated
```{r}

## Filter to treated vs untreated
Treated_untreated_indices <-
    sample_table %>%    
    select(Run_Order)

lipidDDLPS_DMEM <- lipidDDLPS %>%
    select(as.vector(t(Treated_untreated_indices)))

cell <-
    sample_table %>%    
    select(Cell_line)

treatment <-
    sample_table %>%
    select(Treatment_Status)


## return eta_sq(fit)
if(!exists("TreVsUntreLMME")){
    TreVsUntreLMME<-t(pbapply(lipidDDLPS_DMEM, 1, function(x){
        df<-data.frame(y=x,cell=unlist(cell),treatment=unlist(treatment))
        colnames(df)[1]="y"
        treated<-df %>%
            filter(treatment=="Atorvastatin") %>%
            select(y)
        untreated<-df %>%
            filter(treatment=="none") %>%
            select(y)
        log2fc<-mean(as.matrix(treated))-mean(as.matrix(untreated))
        fit.null<-lmer(y ~ (1|cell), data = df,REML=FALSE)
        fit<-lmer(y ~ treatment + (1|cell), data = df,REML=FALSE)
        anovaFit<-anova(fit.null,fit)
        return(c(anovaFit$'Pr(>Chisq)'[2],as.numeric(eta_sq(fit))[2],log2fc))
    }))

    TreVsUntreLMME<-as.data.frame(TreVsUntreLMME)
    colnames(TreVsUntreLMME)<-c("myp","eta_sq","log2fc")
    TreVsUntreLMMEAdj<-TreVsUntreLMME$mypadj<-p.adjust(TreVsUntreLMME$myp,method="fdr")
    TreVsUntreLMME$name<-rownames(TreVsUntreLMME)
}
hist(TreVsUntreLMMEAdj,breaks=100)

```

#### MDM2 Higher vs Lower
```{r}
Untreated_DMEM_indices <-
    sample_table %>%
    filter(Treatment_Status=="none") %>%
    select(Run_Order)

lipidDDLPS_DMEM_untreated <- lipidDDLPS %>%
    select(as.vector(t(Untreated_DMEM_indices)))

cell <-
    sample_table %>%
    filter(Treatment_Status=="none") %>%
    select(Cell_line)

mdm2_status <-
    sample_table %>%
    filter(Treatment_Status=="none") %>%
    select(MDM2_status)

if(!exists("MDM2HivsLoLMME")){
    MDM2HivsLoLMME<-t(pbapply(lipidDDLPS_DMEM_untreated, 1, function(x){
        df<-data.frame(y=x,cell=unlist(cell),mdm2_status=unlist(mdm2_status))
        colnames(df)[1]="y"
        high<-df %>%
            filter(mdm2_status=="high") %>%
            select(y)
        low<-df %>%
            filter(mdm2_status=="low") %>%
            select(y)
        log2fc<-mean(as.matrix(high))-mean(as.matrix(low))
        fit.null<-lmer(y ~ (1|cell), data = df,REML=FALSE)
        fit<-lmer(y ~ mdm2_status + (1|cell), data = df,REML=FALSE)
        anovaFit<-anova(fit.null,fit)
        return(c(anovaFit$'Pr(>Chisq)'[2],as.numeric(eta_sq(fit))[2],log2fc))
    }))

    MDM2HivsLoLMME<-as.data.frame(MDM2HivsLoLMME)
    colnames(MDM2HivsLoLMME)<-c("myp","eta_sq","log2fc")
    MDM2HivsLoLMMEAdj<-MDM2HivsLoLMME$mypadj<-p.adjust(MDM2HivsLoLMME$myp,method="fdr")
    MDM2HivsLoLMME$name<-rownames(MDM2HivsLoLMME)
}
hist(MDM2HivsLoLMMEAdj,breaks=100)

```

#### Atorvastatin Treated MDM2 Higher vs Atorvastatin Treated MDM2 Lower

```{r, Treated MDM2 high vs low}
Treated_DMEM_indices <-
    sample_table %>%
    filter(Treatment_Status=="Atorvastatin") %>%
    select(Run_Order)

lipidDDLPS_DMEM_treated <- lipidDDLPS %>%
    select(as.vector(t(Treated_DMEM_indices)))

cell <-
    sample_table %>%
    filter(Treatment_Status=="Atorvastatin") %>%
    select(Cell_line)

mdm2_status <-
    sample_table %>%
    filter(Treatment_Status=="Atorvastatin") %>%
    select(MDM2_status)

if(!exists("MDM2HivsLoTreatedLMME")){
    MDM2HivsLoTreatedLMME<-t(pbapply(lipidDDLPS_DMEM_treated, 1, function(x){
        df<-data.frame(y=x,cell=unlist(cell),mdm2_status=unlist(mdm2_status))
        colnames(df)[1]="y"
        high<-df %>%
            filter(mdm2_status=="high") %>%
            select(y)
        low<-df %>%
            filter(mdm2_status=="low") %>%
            select(y)
        log2fc<-mean(as.matrix(high))-mean(as.matrix(low))
        fit.null<-lmer(y ~ (1|cell), data = df,REML=FALSE)
        fit<-lmer(y ~ mdm2_status + (1|cell), data = df,REML=FALSE)
        anovaFit<-anova(fit.null,fit)
        return(c(anovaFit$'Pr(>Chisq)'[2],as.numeric(eta_sq(fit))[2],log2fc))
    }))

    MDM2HivsLoTreatedLMME<-as.data.frame(MDM2HivsLoTreatedLMME)
    colnames(MDM2HivsLoTreatedLMME)<-c("myp","eta_sq","log2fc")
    MDM2HivsLoTreatedLMMEAdj<-MDM2HivsLoTreatedLMME$mypadj<-p.adjust(MDM2HivsLoTreatedLMME$myp,method="fdr")
    MDM2HivsLoTreatedLMME$name<-rownames(MDM2HivsLoTreatedLMME)
}
```

### Volcano plots

##### Function

```{r}
volcano_plotter <- function(df,title){
    T_test_results<-df
    
    ## Make insignificant points smaller and more transparent
    T_test_results$significant = abs(T_test_results$log2fc) > fc_thresh &
        -log10(T_test_results$mypadj) > -log10(p_thresh)
    T_test_results$size <- ifelse(T_test_results$significant,4,2)
    
    ## Color code by lipid class
    T_test_results$MainClass<-lipidomic_key$LM.Main.Class[match(T_test_results$name,lipidomic_key$name)]
    palette_names<-as.vector(sort(unique(T_test_results$MainClass)))
    T_test_results$MainClass<-sapply(1:nrow(T_test_results),function(x){
        ifelse(T_test_results$significant[x],
                    return(as.vector(T_test_results$MainClass)[x]),return("Not Significant"))
    })
      
	 legend_labels<-sort(unique(sapply(T_test_results$MainClass, function(x){
             if(!is.na(x)){
		 return(paste0(x," (n=",length(which(T_test_results$MainClass==x)),")"))
             }else{
		 return(x)
             }
	 })))
      
	 ##legend_labels<-legend_labels[-which(is.na(legend_labels))]
      
	 palette<-c(brewer.pal(12,"Paired"),"#40E0D0","violet","grey80")
	 names(palette)<-c(palette_names,"Not Significant")
	 palette<-palette[unique(T_test_results$MainClass)]
                                           #palette<-palette[-which(is.na(palette))]
      
    if(any(T_test_results$significant)){
        g = ggplot(data=T_test_results, aes(x=log2fc, y=-log10(mypadj), colour=MainClass)) +
            geom_point(##alpha=0.7,
                       aes(size=size, text=name)) +
            scale_color_manual(values = palette,
                                        #na.value="grey80",
                               labels=legend_labels,
                               name="LipidMaps Main Class") +
            theme_bw(base_size=20) +
                                        #scale_color_brewer(palette="Paired",na.value="grey80") +
            ggtitle(title) +
            theme(plot.title = element_text(hjust = 0.5)) +
            theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
            ylab("-log10 FDR") +
            geom_hline(yintercept = -log10(p_thresh),lty = 2) +
            geom_vline(xintercept = fc_thresh, lty = 2) +
            geom_vline(xintercept = -fc_thresh, lty = 2) +
            scale_size(range=c(2,4)) +
            guides(size=FALSE,colour = guide_legend(override.aes = list(size=10)))
    }else{
        g = ggplot(data=T_test_results, aes(x=log2fc, y=-log10(mypadj))) +
            geom_point(
                aes(text=name)) +
            scale_color_manual(values = palette) +
            theme_bw(base_size=24) +
            ggtitle(title) +
            theme(plot.title = element_text(hjust = 0.5)) +
            theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
            ylab("-log10 FDR") +
            geom_hline(yintercept = -log10(p_thresh),lty = 2) +
            geom_vline(xintercept = fc_thresh, lty = 2) +
            geom_vline(xintercept = -fc_thresh, lty = 2)
        return(g)
    }
    
    T_test_results_sig<-T_test_results[T_test_results$significant,]
    text_size<-round(min(20,350/nrow(T_test_results_sig)),digits=0)
    
    barplot<-ggplot(T_test_results_sig, aes(x=reorder(name,log2fc), y=log2fc ))+
        theme_bw() +
        geom_bar(stat="identity",colour="black",aes(fill = MainClass)) +
        scale_fill_manual(values = palette) +
        coord_flip() +
        ylab("eta squared") +
        theme(axis.title.y = element_blank(),axis.title.x = element_text(size=12,face="bold"),axis.text.y=element_text(size=text_size),axis.text.x=element_text(size=10)) +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
    
    return(list(g,barplot))
}
```


```{r, fig.height=6}
volcano_plotter(TreVsUntreLMME,"Atorvastatin vs Control")[[1]] + labs(x="log2(Atorvastatin)-log2(Untreated)")
volcano_plotter(MDM2HivsLoLMME,"MDM2 High vs Low")[[1]] + xlim(-3.2,3)
volcano_plotter(MDM2HivsLoTreatedLMME, "Treated Cells MDM2 High vs Low")[[1]] + xlim(-3.2,3)

```

## MDM2 higher cell lines only {.tabset}
### Linear Mixed effect models
#### Atorvastatin Treated vs. Untreated
```{r}
   Treated_untreated_indices <-
       sample_table %>%
       filter(MDM2_status=="high") %>%
       select(Run_Order)

   lipidDDLPS_DMEM <- lipidDDLPS %>%
       select(as.vector(t(Treated_untreated_indices)))

   cell <-
       sample_table %>%
       filter(MDM2_status=="high") %>%
       select(Cell_line)

   treatment <-
       sample_table %>%
       filter(MDM2_status=="high") %>%
       select(Treatment_Status)

   if(!exists("TreVsUntreLMME_high")){
       TreVsUntreLMME_high<-t(pbapply(lipidDDLPS_DMEM, 1, function(x){
           df<-data.frame(y=x,cell=unlist(cell),treatment=unlist(treatment))
           colnames(df)[1]="y"
           treated<-df %>%
               filter(treatment=="Atorvastatin") %>%
               select(y)
           untreated<-df %>%
               filter(treatment=="none") %>%
           select(y)
           log2fc<-mean(as.matrix(treated))-mean(as.matrix(untreated))
           fit.null<-lmer(y ~ (1|cell), data = df,REML=FALSE)
           fit<-lmer(y ~ treatment + (1|cell), data = df,REML=FALSE)
           anovaFit<-anova(fit.null,fit)
           return(c(anovaFit$'Pr(>Chisq)'[2],as.numeric(eta_sq(fit))[2],log2fc))
       }))
       
       TreVsUntreLMME_high<-as.data.frame(TreVsUntreLMME_high)
       colnames(TreVsUntreLMME_high)<-c("myp","eta_sq","log2fc")
       TreVsUntreLMMEAdj_high<-TreVsUntreLMME_high$mypadj<-p.adjust(TreVsUntreLMME_high$myp,method="fdr")
       TreVsUntreLMME_high$name<-rownames(TreVsUntreLMME_high)
   }
   hist(TreVsUntreLMMEAdj_high,breaks=100)

```

### Volcano Plots

```{r, fig.height=6,eval=TRUE}

volcano_plotter(TreVsUntreLMME_high,"Atorvastatin vs Control, MDM2 higher only")[[1]] + xlim(-3.55,3) + ylim(0,10) + labs(x="log2(Atorvastatin)-log2(Untreated)")

```

## MDM2 lower cell lines only {.tabset}
### Linear Mixed effect models
#### Atorvastatin Treated vs. Untreated
```{r}
Treated_untreated_indices <-
       sample_table %>%
       filter(MDM2_status=="low") %>%
       select(Run_Order)

   lipidDDLPS_DMEM <- lipidDDLPS %>%
       select(as.vector(t(Treated_untreated_indices)))

   cell <-
       sample_table %>%
       filter(MDM2_status=="low") %>%
       select(Cell_line)

   treatment <-
       sample_table %>%
       filter(MDM2_status=="low") %>%
       select(Treatment_Status)


if(!exists("TreVsUntreLMME_low")){
   TreVsUntreLMME_low<-t(pbapply(lipidDDLPS_DMEM, 1, function(x){
       df<-data.frame(y=x,cell=unlist(cell),treatment=unlist(treatment))
       colnames(df)[1]="y"
       treated<-df %>%
           filter(treatment=="Atorvastatin") %>%
           select(y)
       untreated<-df %>%
           filter(treatment=="none") %>%
           select(y)
       log2fc<-mean(as.matrix(treated))-mean(as.matrix(untreated))
       fit.null<-lmer(y ~ (1|cell), data = df,REML=FALSE)
       fit<-lmer(y ~ treatment + (1|cell), data = df,REML=FALSE)
       anovaFit<-anova(fit.null,fit)
       return(c(anovaFit$'Pr(>Chisq)'[2],as.numeric(eta_sq(fit))[2],log2fc))
   }))


   TreVsUntreLMME_low<-as.data.frame(TreVsUntreLMME_low)
   colnames(TreVsUntreLMME_low)<-c("myp","eta_sq","log2fc")
   TreVsUntreLMME_lowAdj<-TreVsUntreLMME_low$mypadj<-p.adjust(TreVsUntreLMME_low$myp,method="fdr")
   TreVsUntreLMME_low$name<-rownames(TreVsUntreLMME_low)
   write.csv(TreVsUntreLMME_low,file="~/Desktop/LMME_lipidomic_results.csv")
   }

   hist(TreVsUntreLMME_lowAdj,breaks=100)

```

### Volcano Plots
```{r, Volcano Plots MDM2 Low}

volcano_plotter(TreVsUntreLMME_low,"Atorvastatin vs Control, MDM2 lower only")[[1]] + xlim(-3.5,3.5) + ylim(0,7) + labs(x="log2(Atorvastatin)-log2(Untreated)")

```

