---
title: Evaluating differences between MDM2 higher and lower DDLPS cell lines using data from the Metabolon Platform
author: "Andrew Patt"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
    theme: cerulean
    highlight: tango
---

```{r,echo=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, 
                      fig.height=6,fig.width = 10,fig.fullwidth = TRUE)
##rm(list=ls())
```
The code in this script was used to identify metabolomic differences between MDM2 higher and lower cell lines, as well as differences driven by MDM2 inhibitor treatment. Code for some of the figures in the paper as well as for generating the supplementary data tables is also included.

# Install and load required packages

```{r}
if(!require("tidyverse")){
  install.packages("tidyverse")
}
if(!require("lme4")){
  install.packages("lme4")
}
if(!require("RColorBrewer")){
  install.packages("RColorBrewer")
}
if(!require("cowplot")){
  install.packages("cowplot")
}
if(!require("gplots")){
  install.packages("gplots")
}
if(!require("readxl")){
  install.packages("readxl")
v}
if(!require("reshape")){
  install.packages("reshape")
}
if(!require("sjstats")){
  install.packages("sjstats")
}
if(!require("raster")){
  install.packages("raster")
}
if(!require("RaMP")){
  install.packages("devtools")
  library(devtools)
  install_github("mathelab/RAMP-DB")
}
```

# Load data and metadata, and define statistical groups

```{r}
##rm(list=ls())

load("~/Desktop/Liposarcoma project/SarcomaLines.Rdata")
##load("~/Desktop/Liposarcoma project/final_figures/final_figures_org_file/figureS1D.Rda")
mycateg=gsub("_.*","",sampmeta$CLIENT.IDENTIFIER)
mycol=c("red","blue")

getind <- function(subcateg, allcateg) {
    ind=c()
    for (i in 1:length(subcateg)) {
        ind=c(ind, which(allcateg==subcateg[i]))
    }
    return(ind)
}

firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}

mdm2lo_groups=c("863","815","224B")
mdm2hi_groups=c("224","246","141")
all_groups=c(mdm2hi_groups,mdm2lo_groups)

mdm2lo=getind(subcateg=mdm2lo_groups,allcateg=mycateg)
mdm2hi=getind(subcateg=mdm2hi_groups,allcateg=mycateg)

mdm2lotreated_groups <- c("863RG5","815RG5")
mdm2hitreated_groups <- c("246RG5","141RG5")

mdm2loTreated=getind(subcateg=mdm2lotreated_groups,allcateg=mycateg)
mdm2hiTreated=getind(subcateg=mdm2hitreated_groups,allcateg=mycateg)

mysamps = c(mdm2hi,mdm2lo,mdm2hiTreated,mdm2loTreated)
groups<-unique(mycateg[mysamps])

```

# Filter out metabolites with a high proportion of missing values or a high coefficient of variation

```{r}
metabDDLPS <- metab[,mysamps]
missingValues <- apply(metabDDLPS,1,function(x) return((length(which(x==min(x)))-1)/length(x)))
hist(missingValues,breaks=100,main="Missing Values by Metabolite")

metabDDLPSnoMV <- metabDDLPS[which(missingValues < 0.5),]

cv <- apply(metabDDLPSnoMV,1,cv)
hist(cv,breaks=100,main="Coefficient of variation by Metabolite")

# Cut off metabolites that have a CV more than 150
metabDDLPSnoMVCV <- metabDDLPSnoMV[which(cv < 150),]

boxplot(metabDDLPSnoMVCV,main="Pre-log transformation")
# Log it
metabDDLPSnoMVCVlog <- log(metabDDLPSnoMVCV, base=2)
metabFiltered <- metabDDLPSnoMVCVlog
boxplot(metabFiltered,main="Post-log transformation")

mycell=mycateg[mysamps]
mystatus=sapply(mycell,function(x) if(x %in% mdm2hi_groups){return("MDM2Hi")}
                else if(x %in% mdm2hitreated_groups){return("MDM2HiTreated")}
                else if(x %in% mdm2lo_groups){return("MDM2Lo")}
                else {return("MDM2LoTreated")})
mysamps <- c(which(as.vector(mystatus)=="MDM2Hi"),which(as.vector(mystatus)=="MDM2Lo"))
mycell=mycell[mysamps]
mystatus <- mystatus[mysamps]

sample_table<-sampmeta %>%
    filter(SAMPLE.NAME %in% colnames(metabFiltered)) %>%
    dplyr::select(SAMPLE.NAME,CLIENT.IDENTIFIER) %>%
    mutate(Cell=gsub("_.*","",CLIENT.IDENTIFIER)) %>%
    mutate(MDM2_status=ifelse(Cell %in% mdm2hi_groups | Cell %in% mdm2hitreated_groups,
            "High",
            "Low")) %>%
    mutate(Treatment_status=ifelse(grepl("RG5",CLIENT.IDENTIFIER),
                                   "Treated",
                                   "Untreated"))

```

## PCA to visualize raw data
```{r,fig.width=10}

## Set ggplot default theme
theme_set(theme_classic(base_size = 14))

mycateg=gsub("_.*","",sampmeta$CLIENT.IDENTIFIER)
library(RColorBrewer)
mycol=c("red","blue")

getind <- function(subcateg, allcateg) {
    ind=c()
    for (i in 1:length(subcateg)) {
        ind=c(ind, which(allcateg==subcateg[i]))
    }
    return(ind)
}

mytitle="MDM2 higher and lower cell lines (untreated)"
mypca=prcomp(t(metabFiltered[,mysamps]),center=T,scale=T)
percvar=round((mypca$sdev)^2 / sum(mypca$sdev^2)*100,2)
mydf=data.frame(PC1=mypca$x[,"PC1"],PC2=mypca$x[,"PC2"],PC3=mypca$x[,"PC3"],
                Status=mystatus,Cell=mycell)
mycol=c("red","blue")


p <- ggplot(mydf,aes(PC1,PC2,color=Status,shape=Cell)) +
  geom_point(aes(PC1,PC2,shape=Cell),size=4) +
  scale_color_manual(values=mycol) +
  xlab(paste0("PC1: ",percvar[1],"% variance")) +
  ylab(paste0("PC2: ",percvar[2],"% variance")) +
  theme_bw(base_size=24) +
  ggtitle(mytitle) +
  theme(axis.line = element_line(colour = "black"),
        ##axis.title=element_text(size=12,face="bold"),
        plot.title=element_text(size=24,face="bold"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.key=element_blank())

p


```

# Visualize Metabolon Superpathway of Detected Metabolites

```{r, Panel composition,fig.width=16}
myPalette<-c("#e6194b","#8B008B","#808080","#f58231","#008080","#46f0f0","#3cb44b","mediumblue","gray80")[c(4,1,3,2,5,7,6,8,9)]
names(myPalette)<-c(names(sort(table(metabmeta$SUPER.PATHWAY),decreasing=TRUE)),
                    "Not Significant")

ggplot(metabmeta,aes(x=forcats::fct_infreq(SUPER.PATHWAY),
                     fill=forcats::fct_infreq(SUPER.PATHWAY))) +
    geom_bar() +
    theme_classic(base_size=30) +
    theme(axis.text.x = element_text(angle=45,hjust=1),
          axis.title.y = element_text(size=26),
          legend.title = element_blank()) +
    labs(x="Chemical Superpathway", y="# of Metabolites") +
    scale_fill_manual(values=myPalette)
```

Metabolon data requires minimal preprocessing since abundances are
normalized to an FBS standard already.

Now that data are quality controlled, we perform linear mixed effect
modeling to identify differentially abundant metabolites, controlling
for cell line effects.

# Linear Mixed Effect Modeling

## MDM2 Higher  vs Lower, Untreated
```{r}
##categ = names(mystatus)[mysamps]
##status = mystatus[mysamps]
mdm2_high_untreated_samps<-sample_table %>%
    filter(MDM2_status=="High" & Treatment_status=="Untreated") %>%
    dplyr::select(SAMPLE.NAME) %>%
    as.matrix() %>%
    as.vector()
mdm2_high_untreated_cells<-sample_table %>%
    filter(MDM2_status=="High" & Treatment_status=="Untreated") %>%
    dplyr::select(Cell) %>%
    as.matrix() %>%
    as.vector()
mdm2_low_untreated_samps<-sample_table %>%
    filter(MDM2_status=="Low" & Treatment_status=="Untreated") %>%
    dplyr::select(SAMPLE.NAME) %>%
    as.matrix() %>%
    as.vector()
mdm2_low_untreated_cells<-sample_table %>%
    filter(MDM2_status=="Low" & Treatment_status=="Untreated") %>%
    dplyr::select(Cell) %>%
    as.matrix() %>%
    as.vector()
if(!exists("mdm2HiLoLMME")){
    mdm2HiLoLMME<-sapply(c(1:nrow(metabFiltered)), function(x){
        df=data.frame(t(metabFiltered[x, c(mdm2_high_untreated_samps,
                                           mdm2_low_untreated_samps)]),
                      categ=c(mdm2_high_untreated_cells,mdm2_low_untreated_cells),
                      status=c(rep("MDM2High",length(mdm2_high_untreated_samps)),
                               rep("MDM2Low",length(mdm2_low_untreated_samps))))
        colnames(df)[1]="y"
        fit.null<-lmer(y ~ (1|categ), data = df,REML=FALSE)
        fit<-lmer(y ~ status + (1|categ), data = df,REML=FALSE)
        anovaFit<-anova(fit.null,fit)
        if(x %% 10 ==0){
        }
        log2fc = mean(as.matrix(metabFiltered[x,mdm2_high_untreated_samps])) - mean(as.matrix(metabFiltered[x,mdm2_low_untreated_samps]))
        return(c(anovaFit$'Pr(>Chisq)'[2],log2fc))
    })
    mdm2HiLoLMME <- as.data.frame(t(mdm2HiLoLMME))
    colnames(mdm2HiLoLMME)<-c("myp","log2fc")
    mdm2HiLoLMME <- mdm2HiLoLMME %>% mutate(mypadj=p.adjust(myp,method="fdr"))
    mdm2HiLoLMME <- mdm2HiLoLMME %>% mutate(alpha=ifelse(mypadj<0.05 & abs(log2fc) > 0.75,0.05,0.025))
    mdm2HiLoLMME <- mdm2HiLoLMME %>% mutate(size=ifelse(mypadj<0.05 & abs(log2fc) > 0.75,4,2))
}
hist(mdm2HiLoLMME$mypadj,breaks=100,main="Adjusted p value distribution of MDM2 higher vs lower comparison")

mdm2HiLoLMME$name = rownames(metabFiltered)

resMDM2HiLoLMME = mdm2HiLoLMME %>%
     mutate(class=ifelse(mypadj<0.05 & abs(log2fc) > 0.75,
                         as.character(metabmeta$SUPER.PATHWAY)[match(name,metabmeta$BIOCHEMICAL)],                         
                         "Not Significant"))

##myPalette<-c("#e6194b","#8B008B","#808080","#f58231","gray80","#008080","#46f0f0","mediumblue")

p<-ggplot(data=resMDM2HiLoLMME, aes(x=log2fc, y=-log10(mypadj), colour=class)) +
    geom_point(aes(size=size,text=name)) +
    theme_bw() +
    ##scale_color_manual(values = myPalette[c(1,2,8,4,5,3,6)]) +
    scale_color_manual(values = myPalette) +
    ggtitle("MDM2 Higher vs Lower") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    ylab("-log10 p-value") +
    xlab("log2(MDM2Hi)/log2(MDM2Lo)") +
    geom_hline(yintercept = -log10(0.05),lty = 2) +
    geom_vline(xintercept = 0.75, lty = 2) +
    geom_vline(xintercept = -0.75, lty = 2) +
    scale_alpha(range=c(0.2,0.7)) +
    scale_size(range=c(2,4)) +
    guides(size=FALSE,alpha=FALSE,label=FALSE)
p

```

### Fig S1D

```{r, fig.height=8,eval=FALSE}

boxplotter <- function(analyte) {
    analyteInd <- match(analyte, rownames(metabFiltered))
    BPdf <- data.frame(categ = names(mystatus)[mysamps],
                       value = t(metabFiltered[analyteInd, mysamps]),
                       status = mystatus[mysamps])
    colnames(BPdf) <- c("categ", "log2Abundance","status")
    
    p <- ggplot(BPdf, aes(x = categ, y = log2Abundance)) +
        ## ggtitle(paste0(analyte, ", padj = ", round(resMDM2HiLoLMME$mypadj[analyteInd],
        ##                                            digits = 3), ", \nlog2FC = ",
        ##                round(resMDM2HiLoLMME$log2fc[analyteInd],digits = 3))) +
        ggtitle(firstup(analyte)) +
        ##geom_jitter(width = 0.125, size = 3) +
        scale_color_brewer(type="seq",palette="Set2") +
        facet_grid(.~status,scales="free") +
        geom_boxplot(aes(x=status,y=log2Abundance,fill=status),alpha=0.25) +
        scale_fill_manual(values=mycol) +
        theme_bw() +
        theme(legend.position = "none",
              axis.title.x=element_blank(),
              axis.text.x=element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), text = element_text(size=16))
    return(p)
}

plot_grid(
    boxplotter("glycosyl-N-palmitoyl-sphingosine"),
    boxplotter("trigonelline (N'-methylnicotinate)"),
    boxplotter("N-acetylglycine"),
    boxplotter("adenine"),
    boxplotter("acetylcholine"),
    boxplotter("Ac-Ser-Asp-Lys-Pro-OH"),
    ncol=2)

```

```{r, eval=FALSE}
mdm2HiLo_Sig <- resMDM2HiLoLMME %>% filter(class!="Not Significant")

p <- ggplot(mdm2HiLo_Sig,aes(x=reorder(name,log2fc), y=log2fc, fill = class)) +
    ## ggtitle("MDM2 High vs. Low Cells") +
    geom_bar(stat = "identity", colour = "black") +
    coord_flip() +
    theme_bw() +
    ylab("log2(MDM2 High)/log2(MDM2 Low)") +
    theme(axis.title.y = element_blank(),
          axis.title.x = element_text(size=16,face="bold"),
          axis.text.y=element_text(face="bold",size=18),
          axis.text.x=element_text(size=10),
          legend.text=element_text(size=16),
          legend.title=element_text(size=18)) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    scale_fill_manual(values=myPalette)##[c(1,2,8,4,3,6)]) #+
##theme(legend.position="bottom",legend.direction="vertical")
p
```

## Treated Hi vs Treated Lo

```{r,fig.height=6,eval=TRUE}

mdm2_high_treated_samps<-sample_table %>%
    filter(MDM2_status=="High" & Treatment_status=="Treated") %>%
    dplyr::select(SAMPLE.NAME) %>%
    as.matrix() %>%
    as.vector()
mdm2_low_treated_samps<-sample_table %>%
    filter(MDM2_status=="Low" & Treatment_status=="Treated") %>%
    dplyr::select(SAMPLE.NAME) %>%
    as.matrix() %>%
    as.vector()
mdm2_high_treated_cells<-sample_table %>%
    filter(MDM2_status=="High" & Treatment_status=="Treated") %>%
    dplyr::select(Cell) %>%
    as.matrix() %>%
    as.vector()
mdm2_low_treated_cells<-sample_table %>%
    filter(MDM2_status=="Low" & Treatment_status=="Treated") %>%
    dplyr::select(Cell) %>%
    as.matrix() %>%
    as.vector()

mysamps<-c(mdm2_high_treated_samps,mdm2_low_treated_samps)
categ = c(mdm2_high_treated_cells,mdm2_low_treated_cells)
status = c(rep("High",length(mdm2_high_treated_samps)),rep("Low",length(mdm2_low_treated_samps)))
if(!exists("mdm2TreHiLoLMME")){
    mdm2TreHiLoLMME<-t(sapply(c(1:nrow(metabFiltered)), function(x){
        df=data.frame(t(metabFiltered[x, mysamps]),categ,status)
        colnames(df)[1]="y"
        fit.null<-lmer(y ~ (1|categ), data = df,REML=FALSE)
        fit<-lmer(y ~ status + (1|categ), data = df,REML=FALSE)
        anovaFit<-anova(fit.null,fit)
        log2fc = mean(as.matrix(metabFiltered[x,mdm2_high_treated_samps])) - mean(as.matrix(metabFiltered[x,mdm2_low_treated_samps]))
        return(c(anovaFit$'Pr(>Chisq)'[2],log2fc))
    }))
    
    mdm2TreHiLoLMME <- as.data.frame(mdm2TreHiLoLMME)
    colnames(mdm2TreHiLoLMME)<-c("myp","log2fc")
    mdm2TreHiLoLMME <- mdm2TreHiLoLMME %>% mutate(mypadj=p.adjust(myp,method="fdr"))
    mdm2TreHiLoLMME <- mdm2TreHiLoLMME %>% mutate(alpha=ifelse(mypadj<0.05 & abs(log2fc) > 0.75,0.05,0.025))
    mdm2TreHiLoLMME <- mdm2TreHiLoLMME %>% mutate(size=ifelse(mypadj<0.05 & abs(log2fc) > 0.75,4,2))
}
hist(mdm2TreHiLoLMME[,1],breaks=100)

mdm2TreHiLoLMME = mdm2TreHiLoLMME %>%
    mutate(class=ifelse(mypadj<0.05 & abs(log2fc) > 0.75,
                        as.character(metabmeta$SUPER.PATHWAY[match(rownames(metabFiltered),metabmeta$BIOCHEMICAL)]),
                        "Not Significant"))

mdm2TreHiLoLMME$name = rownames(metabFiltered)

p<-ggplot(data=mdm2TreHiLoLMME, aes(x=log2fc, y=-log10(mypadj), colour=class)) +
    geom_point(aes(size=size,text=name)) +
    theme_classic(base_size=20) +
    scale_color_manual(values = myPalette) +
    ggtitle("MDM2 higher vs lower (treated)") +
    theme(plot.title = element_text(hjust = 0.5)) +
    ylab("-log10 p-value") +
    xlab("log2(MDM2Hi Treated)/log2(MDM2Lo Treated)") +
    geom_hline(yintercept = -log10(0.05),lty = 2) +
    geom_vline(xintercept = 0.75, lty = 2) +
    geom_vline(xintercept = -0.75, lty = 2) +
    scale_alpha(range=c(0.2,0.7)) +
    scale_size(range=c(2,4)) +
    guides(size=FALSE,alpha=FALSE,label=FALSE,
           colour = guide_legend(override.aes = list(size=10)))
##ggplotly(p)
p

##write.csv(mdm2TreHiLoLMME %>% dplyr::select(name,class,log2fc,myp,mypadj), file="~/Desktop/LMME_metab.csv")
```

## Treated Hi vs Untreated Hi

```{r,fig.height=6,eval=TRUE}
mdm2_high_untreated_samps_no224<-sample_table %>%
    filter(MDM2_status=="High" & Treatment_status=="Untreated" & Cell != "224") %>%
    dplyr::select(SAMPLE.NAME) %>%
    as.matrix() %>%
    as.vector()
mdm2_high_untreated_cells_no224<-sample_table %>%
    filter(MDM2_status=="High" & Treatment_status=="Untreated" & Cell != "224") %>%
    dplyr::select(Cell) %>%
    as.matrix() %>%
    as.vector()

mysamps = c(mdm2_high_treated_samps,mdm2_high_untreated_samps_no224)
  categ = c(mdm2_high_treated_cells,mdm2_high_untreated_cells_no224)
  status = c(rep("High_treated",length(mdm2_high_treated_samps)),rep("high",length(mdm2_high_untreated_samps_no224)))
if(!exists("mdm2HiTreLMME")){
  mdm2HiTreLMME<-t(sapply(c(1:nrow(metabFiltered)), function(x){
      df=data.frame(t(metabFiltered[x, mysamps]),categ,status)
      colnames(df)[1]="y"
      fit.null<-lmer(y ~ (1|categ), data = df,REML=FALSE)
      fit<-lmer(y ~ status + (1|categ), data = df,REML=FALSE)
      anovaFit<-anova(fit.null,fit)
      if(x %% 10 ==0){
          #print(x)
          #print(anovaFit$'Pr(>Chisq)'[2])
      }
      log2fc = mean(as.matrix(metabFiltered[x,mdm2_high_treated_samps])) - mean(as.matrix(metabFiltered[x,mdm2_high_untreated_samps]))
      return(c(anovaFit$'Pr(>Chisq)'[2],log2fc))
  }))

  mdm2HiTreLMME <- as.data.frame(mdm2HiTreLMME)
  colnames(mdm2HiTreLMME)<-c("myp","log2fc")
  mdm2HiTreLMME <- mdm2HiTreLMME %>% mutate(mypadj=p.adjust(myp,method="fdr"))
  mdm2HiTreLMME <- mdm2HiTreLMME %>% mutate(alpha=ifelse(mypadj<0.05 & abs(log2fc) > 0.75,0.05,0.025))
  mdm2HiTreLMME <- mdm2HiTreLMME %>% mutate(size=ifelse(mypadj<0.05 & abs(log2fc) > 0.75,4,2))
}
  hist(mdm2HiTreLMME[,1],breaks=100)

mdm2HiTreLMME$class = metabmeta$SUPER.PATHWAY[match(rownames(metabFiltered),metabmeta$BIOCHEMICAL)]
mdm2HiTreLMME$name = rownames(metabFiltered)

  p<-ggplot(data=mdm2HiTreLMME, aes(x=log2fc, y=-log10(mypadj))) +
      geom_point(aes(size=size,text=name),color="gray80") +
      theme_bw() +
      scale_color_manual(values = myPalette) +
      ggtitle("Treated vs Untreated, MDM2 High only") +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
      ylab("-log10 p-value") +
      xlab("log2(MDM2Hi Treated)-log2(MDM2Hi)") +
      geom_hline(yintercept = -log10(0.05),lty = 2) +
      geom_vline(xintercept = 0.75, lty = 2) +
      geom_vline(xintercept = -0.75, lty = 2) +
      scale_alpha(range=c(0.2,0.7)) +
      scale_size(range=c(2,4)) +
      guides(size=FALSE,alpha=FALSE,label=FALSE)
##ggplotly(p)
p

##write.csv(mdm2HiTreLMME %>% dplyr::select(name,class,log2fc,myp,mypadj), file="~/Desktop/LMME_metab.csv")
```

## Treated Lo vs Lo

Next comparing 815 and 863 treated and untreated (224B is filtered out since there is no treatment data available for this cell line)

```{r,fig.height=6}
mdm2_low_untreated_samps_no224B<-sample_table %>%
    filter(MDM2_status=="Low" & Treatment_status=="Untreated" & Cell != "224B") %>%
    dplyr::select(SAMPLE.NAME) %>%
    as.matrix() %>%
    as.vector()
mdm2_low_untreated_cells_no224B<-sample_table %>%
    filter(MDM2_status=="Low" & Treatment_status=="Untreated" & Cell != "224B") %>%
    dplyr::select(Cell) %>%
    as.matrix() %>%
    as.vector()

mysamps = c(mdm2_low_treated_samps,mdm2_low_untreated_samps_no224B)
categ = c(mdm2_low_treated_cells,mdm2_low_untreated_cells_no224B)
status = c(rep("Low_treated",length(mdm2_low_treated_samps)),
           rep("Low",length(mdm2_low_untreated_samps_no224B)))

if(!exists("mdm2LoTreLMME")){
  mdm2LoTreLMME<-t(sapply(c(1:nrow(metabFiltered)), function(x){
      df=data.frame(t(metabFiltered[x, mysamps]),categ,status)
      colnames(df)[1]="y"
      fit.null<-lmer(y ~ (1|categ), data = df,REML=FALSE)
      fit<-lmer(y ~ status + (1|categ), data = df,REML=FALSE)
      anovaFit<-anova(fit.null,fit)
      log2fc = mean(as.matrix(metabFiltered[x,mdm2_low_treated_samps])) - mean(as.matrix(metabFiltered[x,mdm2lo]))
      return(c(anovaFit$'Pr(>Chisq)'[2],log2fc))
  }))

  mdm2LoTreLMME <- as.data.frame(mdm2LoTreLMME)
  colnames(mdm2LoTreLMME)<-c("myp","log2fc")
  mdm2LoTreLMME <- mdm2LoTreLMME %>% mutate(mypadj=p.adjust(myp,method="fdr"))
  mdm2LoTreLMME <- mdm2LoTreLMME %>% mutate(alpha=ifelse(mypadj<0.05 & abs(log2fc) > 0.75,0.05,0.025))
  mdm2LoTreLMME <- mdm2LoTreLMME %>% mutate(size=ifelse(mypadj<0.05 & abs(log2fc) > 0.75,4,2))
  mdm2LoTreLMME$name = rownames(metabFiltered)
}
hist(mdm2LoTreLMME$myp,breaks=100)


mdm2LoTreLMME = mdm2LoTreLMME %>%
    mutate(class=ifelse(mypadj<0.1 & abs(log2fc) > 0.75,
                        as.character(metabmeta$SUPER.PATHWAY[match(name,metabmeta$BIOCHEMICAL)]),
                        "Not Significant"))

p<-ggplot(data=mdm2LoTreLMME, aes(x=log2fc, y=-log10(mypadj), colour=class)) +
    geom_point(aes(size=size,text=name)) +
    theme_bw() +
    scale_color_manual(values = myPalette) +
    ggtitle("Treated vs Untreated, MDM2 low only") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    ylab("-log10 p-value") +
    xlab("log2(MDM2Lo Treated - MDM2Lo)") +
    geom_hline(yintercept = -log10(0.1),lty = 2) +
    geom_vline(xintercept = 0.75, lty = 2) +
    geom_vline(xintercept = -0.75, lty = 2) +
    scale_alpha(range=c(0.2,0.7)) +
    scale_size(range=c(2,4)) +
    guides(size=FALSE,alpha=FALSE,label=FALSE)

##ggplotly(p)
p

##write.csv(mdm2LoTreLMME %>% dplyr::select(name,class,log2fc,myp,mypadj), file="~/Desktop/LMME_metab.csv")

```
### Fig 2D

```{r}
mdm2Lo_Treated_vs_UntreatedSig <- mdm2LoTreLMME %>% filter(class!="Not Significant")

fig2D <- ggplot(mdm2Lo_Treated_vs_UntreatedSig,aes(x=reorder(name,log2fc), y=log2fc, fill = class)) +
    ggtitle("MDM2 lower cells") +
    geom_bar(stat = "identity", colour = "black") +
    coord_flip() +
    theme_bw(base_size=20) +
    ylab("log2(Treated)/log2(Untreated)") +
    theme(axis.title.y = element_blank(),axis.title.x = element_text(size=12,face="bold"),axis.text.y=element_text(face="bold")) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    scale_fill_manual(values=myPalette) #+
##theme(legend.position="bottom",legend.direction="vertical")
fig2D

```

## Treated vs Untreated

```{r,fig.height=6,eval=FALSE}
mysamps = c(mdm2_high_treated_samps,mdm2_low_treated_samps,mdm2hi,mdm2lo)
  categ = c(mycateg[mdm2hiTreated],mycateg[mdm2loTreated],mycateg[mdm2hi],mycateg[mdm2lo])
  status = c(rep("Treated",length(mdm2_high_treated_samps)+length(mdm2_low_treated_samps)),rep("Untreated",length(mdm2lo)+length(mdm2lo)))
if(!exists("mdm2TrevsUntre")){
  mdm2TrevsUntre<-t(sapply(c(1:nrow(metabFiltered)), function(x){
      df=data.frame(t(metabFiltered[x, mysamps]),categ,status)
      colnames(df)[1]="y"
      fit.null<-lmer(y ~ (1|categ), data = df,REML=FALSE)
      fit<-lmer(y ~ status + (1|categ), data = df,REML=FALSE)
      anovaFit<-anova(fit.null,fit)
      if(x %% 10 ==0){
          #print(x)
          #print(anovaFit$'Pr(>Chisq)'[2])
      }
      log2fc = mean(as.matrix(metabFiltered[x,mdm2_low_treated_samps])) - mean(as.matrix(metabFiltered[x,mdm2lo]))
      return(c(anovaFit$'Pr(>Chisq)'[2],log2fc))
  }))

  mdm2TrevsUntreLMME <- as.data.frame(mdm2TrevsUntre)
  colnames(mdm2TrevsUntreLMME)<-c("myp","log2fc")
  mdm2TrevsUntreLMME <- mdm2TrevsUntreLMME %>% mutate(mypadj=p.adjust(myp,method="fdr"))
  mdm2TrevsUntreLMME <- mdm2TrevsUntreLMME %>% mutate(alpha=ifelse(mypadj<0.05 & abs(log2fc) > 0.75,0.05,0.025))
  mdm2TrevsUntreLMME <- mdm2TrevsUntreLMME %>% mutate(size=ifelse(mypadj<0.05 & abs(log2fc) > 0.75,4,2))
}
hist(mdm2TrevsUntreLMME[,1],breaks=100)

##mdm2TrevsUntreLMME$class = metabmeta$SUPER.PATHWAY[match(rownames(metabFiltered),metabmeta$BIOCHEMICAL)]
mdm2TrevsUntreLMME$class = "Not Significant"
mdm2TrevsUntreLMME$name = rownames(metabFiltered)

p<-ggplot(data=mdm2TrevsUntreLMME, aes(x=log2fc, y=-log10(mypadj))) +
    geom_point(aes(size=size,text=name),colour="gray80") +
    theme_bw() +
    ##scale_color_manual(values = myPalette) +
    ggtitle("Treated vs Untreated") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    ylab("-log10 p-value") +
    xlab("log2(Treated)-log2(Untreated)") +
    geom_hline(yintercept = -log10(0.05),lty = 2) +
    geom_vline(xintercept = 0.75, lty = 2) +
    geom_vline(xintercept = -0.75, lty = 2) +
    scale_alpha(range=c(0.2,0.7)) +
    scale_size(range=c(2,4)) +
    guides(size=FALSE,alpha=FALSE,label=FALSE,colour=FALSE)
##ggplotly(p)
p

```

# Post-hoc Analyses

## Hierarchical clustering
```{r,eval=FALSE}
metabs_of_interest<-resMDM2HiLoLMME %>%
    filter(class!="Not Significant") %>%
    dplyr::select(name) %>%
    as.matrix() %>%
    ##t() %>%
    as.vector()

metabs_of_interest<-unique(c("adenine",metabs_of_interest,as.matrix(mdm2Lo_Treated_vs_UntreatedSig$name)))




metab_hclust <- metabFiltered_withRG1[metabs_of_interest,
                              c(which(mystatus_filtered=="MDM2Hi"),
                         which(mystatus_filtered=="MDM2Lo"),
                         which(mystatus_filtered=="MDM2HiRG5"),
                         which(mystatus_filtered=="MDM2LoRG5"))]
labels <- sampmeta_filtered %>%
    filter(SAMPLE.NAME %in% colnames(metab_hclust)) %>%
    dplyr::select(Group)

labels <- make.names(as.vector(t(as.matrix(labels))),unique=T)
labels=gsub("X","",labels)



library(gplots)
heatmap.2(as.matrix(metab_hclust), scale = "none", col = colorRampPalette(c("blue", "white","red"))(100), 
          trace = "none", density.info = "none",
          margins = c(5, 15),
          ColSideColors = c(rep("red", 13), rep("blue", 13),
                            rep("orange", 6),
                            rep("green", 6)),
          RowSideColors = c("yellow",rep("forestgreen",17),rep("purple",17)),
          labRow=rownames(metab_hclust),
          labCol=labels,
          Rowv=FALSE)

metabs_of_interest_filtered<-c("glycosyl-N-palmitoyl-sphingosine",
                               "ornithine",
                               ##"4-hydroxyglutamate",
                               ##"palmitoylcholine",
                               "adenine",
                               "1,2-dioleoyl-GPC (18:1/18:1)*",      
                               "1-stearoyl-2-arachidonoyl-GPE (18:0/20:4)",
                               "1-stearoyl-2-oleoyl-GPE (18:0/18:1)",
                               "glutamate",
                               "myristate")

metab_hclust <- metabFiltered_withRG1[metabs_of_interest_filtered,
                              c(which(mystatus_filtered=="MDM2Hi"),
                         which(mystatus_filtered=="MDM2Lo"),
                         which(mystatus_filtered=="MDM2HiRG5"),
                         which(mystatus_filtered=="MDM2LoRG5"))]
labels <- sampmeta_filtered %>%
    filter(SAMPLE.NAME %in% colnames(metab_hclust)) %>%
    dplyr::select(Group)

labels <- make.names(as.vector(t(as.matrix(labels))),unique=T)
labels=gsub("X","",labels)

heatmap.2(as.matrix(metab_hclust), scale = "none", col = colorRampPalette(c("blue", "white","red"))(100), 
          trace = "none", density.info = "none",
          margins = c(5, 15),
          ColSideColors = c(rep("red", 13), rep("blue", 13),
                            rep("orange", 6),
                            rep("green", 6)),
          #RowSideColors = c("yellow",rep("forestgreen",17),rep("purple",17)),
          labRow=rownames(metab_hclust),
          labCol=labels,
          Rowv=FALSE)

metab_hclust<-cor(metab_hclust)
##metab_hclust<-1-metab_hclust

heatmap.2(metab_hclust, scale = "none", col = bluered(100), 
          trace = "none", density.info = "none",
          ColSideColors = c(rep("red", 13), rep("blue", 13),
                            rep("orange", 6),
                            rep("forestgreen", 6)),
          labRow=labels,
          labCol=labels)

```

## Trends for Glycosyl-N-Palmitoyl-Sphingosine
```{r, glycosyl-n-palmitoyl-sphingosine,eval=FALSE}

sampmeta_filtered<-left_join(data.frame(SAMPLE.NAME=colnames(metabFiltered)),sampmeta, by = "SAMPLE.NAME")

mycateg_filtered=gsub("_.*","",sampmeta_filtered$CLIENT.IDENTIFIER)

mdm2loTreatedRG1_groups = c("863RG1","815RG1")
mdm2hiTreatedRG1_groups = c("246RG1")

mdm2hiTreatedRG1_filtered = getind(subcateg=mdm2hiTreatedRG1_groups,allcateg=mycateg_filtered)
mdm2loTreatedRG1_filtered = getind(subcateg=mdm2loTreatedRG1_groups,allcateg=mycateg_filtered)

mdm2lo_groups=c("863","815","224B")
mdm2hi_groups=c("224","246","141")
all_groups=c(mdm2hi_groups,mdm2lo_groups)

mdm2lo_filtered=getind(subcateg=mdm2lo_groups,allcateg=mycateg_filtered)
mdm2hi_filtered=getind(subcateg=mdm2hi_groups,allcateg=mycateg_filtered)

mdm2lotreated_groups <- c("863RG5","815RG5")
mdm2hitreated_groups <- c("246RG5","141RG5")

mdm2loTreated_filtered=getind(subcateg=mdm2lotreated_groups,allcateg=mycateg_filtered)
mdm2hiTreated_filtered=getind(subcateg=mdm2hitreated_groups,allcateg=mycateg_filtered)

mystatus_filtered=sapply(as.vector(sampmeta_filtered$MANIFEST.ID),function(x) if(x %in% mdm2hi_groups){return("MDM2Hi")}
                else if(x %in% mdm2hitreated_groups){return("MDM2HiRG5")}
                else if(x %in% mdm2lo_groups){return("MDM2Lo")}
                else if(x %in% mdm2lotreated_groups){return("MDM2LoRG5")}
                else if(x %in% mdm2hiTreatedRG1_groups){return("MDM2HiRG1")}
                else if(x %in% mdm2loTreatedRG1_groups){return("MDM2LoRG1")})

boxplotter <- function(analyte="glycosyl-N-palmitoyl-sphingosine",group1,group2){
    analyteInd <- match(analyte, rownames(metabFiltered))
    BPdf <- data.frame(categ = names(mystatus_filtered)[c(group1,group2)],
                       value = t(metabFiltered[analyteInd, c(group1,group2)]),
                       status = factor(mystatus_filtered[c(group1,group2)],
                                       levels=unique(mystatus_filtered[c(group1,group2)])))
    colnames(BPdf) <- c("categ", "log2Abundance","status")
    
    p <- ggplot(BPdf, aes(x = categ, y = log2Abundance)) +
        geom_jitter(width = 0.125, size = 3) +
        scale_color_brewer(type="seq",palette="Set2") +
        facet_grid(.~status,scales="free") +
        geom_boxplot(aes(x=status,y=log2Abundance),alpha=0.25,fill="gray40") +
        scale_fill_manual(values=mycol) +
        theme_bw() +
        ggtitle(analyte) +
        theme(legend.position = "none",axis.title.x=element_blank(), panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(), text = element_text(size=18),
              axis.text.x = element_text(angle=45,hjust=1))
    return(p)
}

boxplotter("glycosyl-N-palmitoyl-sphingosine",mdm2lo_filtered,mdm2hi_filtered)
boxplotter("glycosyl-N-palmitoyl-sphingosine",mdm2hi_filtered,c(mdm2hiTreatedRG1_filtered,mdm2hiTreated_filtered))
boxplotter("glycosyl-N-palmitoyl-sphingosine",mdm2lo_filtered,c(mdm2loTreatedRG1_filtered,mdm2loTreated_filtered))

boxplotter("glycosyl-N-palmitoyl-sphingosine",mdm2hi_filtered,mdm2loTreated_filtered)
boxplotter("glycosyl-N-stearoyl-sphingosine",mdm2hi_filtered,mdm2loTreated_filtered)
boxplotter("N-palmitoyl-sphingosine (d18:1/16:0)",mdm2hi_filtered,mdm2loTreated_filtered)
boxplotter("N-palmitoyl-sphingosine (d18:1/16:0)",mdm2lo_filtered,mdm2loTreated_filtered)
```
# Save final dataframe and metadata

```{r, Save final dataframe and metadata,echo=FALSE,eval=FALSE}
saveRDS(metabFiltered,"~/Desktop/Liposarcoma project/final_dataframes_and_metadata/metab_data.rds")

metab_metadata<-list(mdm2hi_filtered,
                 mdm2lo_filtered,
                 mdm2hiTreated_filtered,
                 mdm2loTreated_filtered,
                 mdm2hiTreatedRG1_filtered,
                 mdm2loTreatedRG1_filtered,
                 left_join(data.frame(SAMPLE.NAME=colnames(metabFiltered)),sampmeta,by="SAMPLE.NAME") %>% dplyr::select(MANIFEST.ID)
)

names(metab_metadata)<-c("mdm2hiUntreatedSamps",
                         "mdm2loUntreatedSamps",
                         "mdm2hiRG5Samps",
                         "mdm2loRG5Samps",
                         "mdm2hiRG1Samps",
                         "mdm2loRG1Samps",
                         "cell_lines")
saveRDS(metab_metadata,"~/Desktop/Liposarcoma project/final_dataframes_and_metadata/metab_metadata.rds")

```

